---
title: "Case Study: Customer Retention and Acquisition"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Install required packages if you need them and load
```{r, echo=FALSE}
rm(list=ls()) # clear environment
#install.packages("SMCRM") # maybe function to install if we don't have it?

library(SMCRM)
library(dplyr)
library(randomForest)
library(caret)
library(ggplot2)
library(pdp)
```

## Get the data
```{r, echo=FALSE}
data("acquisitionRetention")
df <- acquisitionRetention
str(df)
summary(df)

# Convert binary variables from numeric to factors
df$acquisition <- as.factor(df$acquisition)
df$industry <- as.factor(df$industry)

# Drop customer variable
df = select(df, -customer)
```

## Quick checks
```{r, echo=FALSE}
print(paste("Total rows:", count(df)))
print(paste("Missing rows:", count(df) - count(na.omit(df)))) #no missing data

# If duration is 0 acquisition is 0; if duration > 0 acquisition is 1. Predicting duration predicts acquisition.
check_acquisition_duration <- df %>%
  filter((acquisition == 1 & duration > 0) | (acquisition == 0 & duration == 0))
print(paste("Acquisition calculated from Duration:", count(df) == count(check_acquisition_duration)))

# Check balance
df %>% count(acquisition) #0:162 ; 1:338
```

## Split the data
```{r, echo=FALSE}
set.seed(123)
trainIndex <- createDataPartition(df$acquisition, p = 0.7, list = FALSE)
train <- df[trainIndex, ]
test <- df[-trainIndex, ]

# Since we know acquisition based on the value of duration, we can't include this variable as an independent variable
independent_vars <- setdiff(names(train), c("duration", "acquisition"))
```


## Build Random Forest Model
```{r, echo=FALSE}
rf_duration <- randomForest(duration ~ ., 
                            data = train[, c(independent_vars, "duration")], 
                            importance = TRUE, 
                            ntree = 500)

# Predict Duration on the test set
rf_duration_preds <- predict(rf_duration, test[, independent_vars])

# Evaluate Duration predictions
regression_metrics <- data.frame(
  RMSE = sqrt(mean((rf_duration_preds - test$duration)^2)),
  R2 = cor(rf_duration_preds, test$duration)^2
)
print("Regression Metrics for Duration Prediction:")
print(regression_metrics)


# Derive Acquisition predictions
rf_acquisition_preds <- ifelse(rf_duration_preds > 0, 1, 0)

# Evaluate Acquisition predictions
confusion_matrix <- confusionMatrix(factor(rf_acquisition_preds), factor(test$acquisition))
print("Confusion Matrix for Acquisition Prediction:")
print(confusion_matrix)

#Feature Importance
importance <- importance(rf_duration)
var_importance <- data.frame(Variable = rownames(importance),
                             Importance = importance[, 1])
var_importance <- var_importance[order(var_importance$Importance, decreasing = TRUE), ]
print("Best Variables for Duration Prediction (Best to Worst):")
print(var_importance, 10)

```


## Partial Dependency Plots
```{r, echo=FALSE}
# PDP for the top 3 variables
top_vars <- var_importance$Variable[1:3]
for (var in top_vars) {
  pd <- partial(rf_duration, pred.var = var, train = train[, independent_vars])
  plotPartial(pd, main = paste("Partial Dependence Plot for", var))
}
```
